---
title: "Correlation Analysis of Phenols and N availability"
author: "Zhang Zhenglin"
output:
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
# https://bookdown.org/ansellbr/WEHI_tidyR_course_book/
library(stringr) 
library(data.table)
library(GGally)
library(formatR)
library(readxl)
library(mgcv)
library(writexl)
```

# Data Organisation

## Read from excel

```{r}
preplant <- read_excel("Cor_Base.xlsx", sheet = 1)
str(preplant)

topdress <- read_excel("Cor_Base.xlsx", sheet = 2)
str(topdress)

#cor_data_preplant <- preplant%>%select(-Field,-Blk_Site,-Blk,-Year)
#cor_data_topdress <- topdress%>%select(-Field,-Blk_Site,-Blk,-Year)

cor_data_preplant <- preplant%>%select(-Blk_Site,-Blk,-Year)
cor_data_topdress <- topdress%>%select(-Blk_Site,-Blk,-Year)
```

# Correlation matrix (with raw and absolute values)

# Delinated by field

```{r}
preplant_cor_matrix <-ggpairs(cor_data_preplant, aes(col=Field, alpha=0.5))+
    ggtitle("(a) Absolute_Preplant") +
    theme(plot.title = element_text(size = 30, face = "bold"))

topdress_cor_matrix <-ggpairs(cor_data_topdress, aes(col=Field, alpha=0.5))+
    ggtitle("(c) Absolute_Topdress") +
    theme(plot.title = element_text(size = 30, face = "bold"))


#ggsave(preplant_cor_matrix, filename = "huehue_preplant.pdf", height = 30, width = 30, units = "cm")
#ggsave(topdress_cor_matrix, filename = "huehue_topdress.pdf", height = 30, width = 30, units = "cm")
```

# all together

```{r}
all_preplant_cor_matrix <-ggpairs(cor_data_preplant %>% select(-Field, -yield_Mgha), 
                                  upper = list(continuous = wrap("cor", size = 5.5))
                                  )+
                          ggtitle("(a) Preplant") +
                          theme(plot.title = element_text(size = 30, face = "bold"))


all_topdress_cor_matrix <-ggpairs(cor_data_topdress %>% select(-Field, -yield_Mgha), 
                                  upper = list(continuous = wrap("cor", size = 5.5))
                                  )+
                          ggtitle("(b) Topdress") +
                          theme(plot.title = element_text(size = 30, face = "bold"))
                        

ggsave(all_preplant_cor_matrix, filename = "all_preplant.png", height = 34, width = 33, units = "cm", dpi = 400)
ggsave(all_topdress_cor_matrix, filename = "all_topdress.png", height = 34, width = 33, units = "cm", dpi = 400)
```

# Correlation matrix (with differences, obtained betweent treatments by blk)

## Obtain differences by blk and year

## For preplant


difference_preplant <- preplant %>%
  group_by(Year, Blk) %>%                # Group by Year and Blk
  filter(all(c("CR", "RF") %in% Field)) %>%  # Ensure both "CR" and "RF" exist
  summarise(
    TotalP = TotalP[Field == "CR"] - TotalP[Field == "RF"],
    TotalIV = TotalV[Field == "CR"] - TotalV[Field == "RF"],
    TotalC = TotalC[Field == "CR"] - TotalC[Field == "RF"],
    TotalS = TotalS[Field == "CR"] - TotalS[Field == "RF"],
    Total_phenols = Total_phenols[Field == "CR"] - Total_phenols[Field == "RF"],
    crop_fertiliser_N = crop_fertiliser_N[Field == "CR"] - crop_fertiliser_N[Field == "RF"],
    crop_soil_N = crop_soil_N[Field == "CR"] - crop_soil_N[Field == "RF"],
    soil_tracer_N = soil_tracer_N[Field == "CR"] - soil_tracer_N[Field == "RF"],
    yield = yield[Field == "CR"] - yield[Field == "RF"]
  ) %>%
  ungroup() 


## For topdress


difference_topdress <- topdress %>%
  group_by(Year, Blk) %>%                # Group by Year and Blk
  filter(all(c("CR", "RF") %in% Field)) %>%  # Ensure both "CR" and "RF" exist
  summarise(
    TotalP = TotalP[Field == "CR"] - TotalP[Field == "RF"],
    TotalIV = TotalV[Field == "CR"] - TotalV[Field == "RF"],
    TotalC = TotalC[Field == "CR"] - TotalC[Field == "RF"],
    TotalS = TotalS[Field == "CR"] - TotalS[Field == "RF"],
    Total_phenols = Total_phenols[Field == "CR"] - Total_phenols[Field == "RF"],
    crop_fertiliser_N = crop_fertiliser_N[Field == "CR"] - crop_fertiliser_N[Field == "RF"],
    crop_soil_N = crop_soil_N[Field == "CR"] - crop_soil_N[Field == "RF"],
    soil_tracer_N = soil_tracer_N[Field == "CR"] - soil_tracer_N[Field == "RF"],
    yield = yield[Field == "CR"] - yield[Field == "RF"]
  ) %>%
  ungroup()

## Remove unneeded columns


cor_data_difference_preplant <- difference_preplant%>%select(-Blk,-Year)
cor_data_difference_topdress <- difference_topdress%>%select(-Blk,-Year)


## all together


all_difference_preplant_cor_matrix <-ggpairs(cor_data_difference_preplant) + 
  ggtitle("(b) Difference_Preplant") +
  theme(plot.title = element_text(size = 30, face = "bold"))


all_difference_topdress_cor_matrix <-ggpairs(cor_data_difference_topdress) + 
  ggtitle("(d) Difference_Topdress")+
  theme(plot.title = element_text(size = 30, face = "bold"))

#ggsave(all_difference_preplant_cor_matrix, filename = "all_difference_preplant.pdf", height = 30, width = 30, units = "cm")
#ggsave(all_difference_topdress_cor_matrix, filename = "all_difference_topdress.pdf", height = 30, width = 30, units = "cm")

```{r}
citation("GGally")
```
