---
title: "Yield Analysis"
author: "Zhang Zhenglin"
output:
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
# https://bookdown.org/ansellbr/WEHI_tidyR_course_book/
library(stringr) 
library(data.table)
library(GGally)
library(formatR)
library(readxl)
```

# Data Organisation

## Read from excel

```{r}
yield <- read_excel("Yield_15N.xlsx", sheet = 1)
str(yield)
```

## Change variable type

```{r}
yield$Plot <- as.factor(yield$Plot)
yield$Field <- as.factor(yield$Field)
yield$Blk <- as.factor(yield$Blk)
yield$TopDress <- as.factor(yield$TopDress)
yield$Year <- as.factor(yield$Year)

str(yield)
```

## Sub dataset for "preplant" and "topdress"

```{r}
preplant <- yield %>% filter(TopDress == "N")
str(preplant)

topdress <- yield %>% filter(TopDress == "Y")
str(topdress)
```

# Initial Visualisation

```{r}
#preplant
ggplot(preplant, aes(y=Yield, x=Field, color=Year)) + geom_boxplot()

#topdress
ggplot(topdress, aes(y=Yield, x=Field, color=Year)) + geom_boxplot()
```

# Preplant

## Comparision by treatment and year

```{r}
preplant_model <- lmer(Yield~Field*Year+(1|Blk), data=preplant)
anova(preplant_model)
```

> All terms significant.

```{r}
preplant_means <- emmeans(preplant_model, spec ='Field', by = 'Year')
preplant_effects <- contrast(preplant_means, method = 'pairwise', adjust = "tukey")
summary(preplant_effects)
cld(preplant_means)
```

> Yield higher for fallow rice in both 2021 and 2022

## Graphing

Create dataframe for error bars

```{r}
preplant_graphing <- preplant %>% group_by(Field, Year) %>% 
   mutate(yield_sd = sd(Yield)) %>% 
  summarise(Yield = mean(Yield),
            yield_sd = mean(yield_sd))
```

```{r}
preplant_yield_graph <-
  ggplot(preplant_graphing, aes(x = Year, y = Yield, fill = Field)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Year", y = "Yield", fill = "Field") +
  scale_fill_manual(values = c("#0072B2","#FFCC66"),labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  geom_errorbar(aes(ymin=Yield-yield_sd, ymax=Yield+yield_sd), width=.2,position=position_dodge(0.7))+
  scale_y_continuous(name=expression("Yield (Mg ha"^{-1}*")"), limits = c(0, 20), expand = c(0, 0))+
  scale_x_discrete(name="")+
  theme_classic()+
  theme(axis.text = element_text(size = 12), axis.title = element_text(size=14))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  annotate(
  "text",
  x = c(1,2),  # X-axis positions for annotations
  y = c(15.7,13),  # Y-axis positions for annotations
  label = "*",
  size = 12,
  vjust = 0  # Adjust vertical position of asterisks
)+
  annotate(
  "text",
  x = c(1.45),  # X-axis positions for annotations
  y = c(19.1),  # Y-axis positions for annotations
  label = "Preplant",
  size = 7,
  vjust = 0  
)

preplant_yield_graph
```

```{r}
ggsave(preplant_yield_graph, filename = "preplant_yield_graph.png", height = 15, width = 20, units = "cm", dpi=1000)
```

# TopDress

## Comparison by treatment

```{r}
topdress_model <- lmer(Yield~Field*Year+(1|Blk), data=topdress)
anova(topdress_model)
```

> Field effect strongest. Year effect present. Likely some Field:Year interation. 

```{r}
topdress_means <- emmeans(topdress_model, spec ='Field', by = 'Year')
topdress_effects <- contrast(topdress_means, method = 'pairwise', adjust = "tukey")
summary(topdress_effects)
cld(topdress_means)
```

> No difference in 2021, but different in 2022.

## Graphing

Create graphing dataframe

```{r}
topdress_graphing <- topdress %>% group_by(Field, Year) %>% 
   mutate(yield_sd = sd(Yield)) %>% 
  summarise(Yield = mean(Yield),
            yield_sd = mean(yield_sd))
```

```{r}
topdress_yield_graph <-
  ggplot(topdress_graphing, aes(x = Year, y = Yield, fill = Field)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(x = "Year", y = "Yield", fill = "Field") +
  scale_fill_manual(values = c("#0072B2","#FFCC66"),labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  geom_errorbar(aes(ymin=Yield-yield_sd, ymax=Yield+yield_sd), width=.2,position=position_dodge(0.7))+
  scale_y_continuous(name=expression("Yield (Mg ha"^{-1}*")"), limits = c(0, 20), expand = c(0, 0))+
  scale_x_discrete(name="")+
  theme_classic()+
  theme(axis.text = element_text(size = 12), axis.title = element_text(size=14))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  annotate(
  "text",
  x = c(2),  # X-axis positions for annotations
  y = c(11),  # Y-axis positions for annotations
  label = "*",
  size = 12,
  vjust = 0  # Adjust vertical position of asterisks
)+
  annotate(
  "text",
  x = c(1.45),  # X-axis positions for annotations
  y = c(19.1),  # Y-axis positions for annotations
  label = "Topdress",
  size = 7,
  vjust = 0  
)

topdress_yield_graph
```


```{r}
ggsave(topdress_yield_graph, filename = "topdress_yield_graph.png", height = 15, width = 20, units = "cm", dpi=1000)
```

