---
title: "Soil N and Fert N Analysis"
author: "Zhang Zhenglin"
output:
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---

# Necessary libraries

```{r message=FALSE,warning=FALSE}
library(knitr)
library(ggplot2)
theme_set(theme_bw())
library(emmeans)
library(multcomp)
library(PLS205)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(Rmisc) 
library(dplyr) #https://r4ds.had.co.nz/ (Chapter 3, Chapter 5, look at filter and select)
# https://bookdown.org/ansellbr/WEHI_tidyR_course_book/
library(stringr) 
library(data.table)
library(GGally)
library(formatR)
library(readxl)
library(mgcv)
library(writexl)

```

# Data Organisation

## Read from excel

```{r}
all_data <- read_excel("15N_N_Uptake_MaturitySummed.xlsx", sheet = 1)
str(all_data)
```

## Clean up variables

```{r}
all_data <- mutate_if(all_data, is.character, as.factor)
all_data$Blk <- as.factor(all_data$Blk)
all_data$Year <- as.factor(all_data$Year)

str(all_data)
```

## Sub dataset for "preplant" and "topdress"

```{r}
preplant <- all_data %>% filter(Topdress == "N")
str(preplant)

topdress <- all_data %>% filter(Topdress == "Y")
str(topdress)
```


## Intial visualisation

```{r}
#preplant soil N
ggplot(preplant, aes(x = Days, y = soil_N, color = Field)) +
  geom_smooth(method = "lm") +
  labs(x = "Days", y = "Soil N uptake", color = "Field", linetype = "Year") +
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_minimal()

#preplant fert N
ggplot(preplant, aes(x = Days, y = fertiliser_N, color = Field)) +
  geom_smooth(method = "lm") +
  labs(x = "Days", y = "Soil N uptake", color = "Field", linetype = "Year") +
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_minimal()

#topdress soil N
ggplot(topdress, aes(x = Days, y = soil_N, color = Field)) +
  geom_smooth(method = "lm") +
  labs(x = "Days", y = "Soil N uptake", color = "Field", linetype = "Year") +
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_minimal()

#topdress fert N
ggplot(topdress, aes(x = Days, y = fertiliser_N, color = Field)) +
  geom_smooth(method = "lm") +
  labs(x = "Days", y = "Soil N uptake", color = "Field", linetype = "Year") +
  scale_linetype_manual(values = c("solid", "dashed")) +
  theme_minimal()
```

# Preplant Soil N

## Model selection and testing

```{r}
preplant_soil_N_model <- lmer(soil_N~Field*Year*Stage+(1|Blk)+(1|Blk:Field), data = preplant)
anova(preplant_soil_N_model)
```

> Strong effect of field. 

```{r}
preplant_soil_N_means = emmeans(preplant_soil_N_model,spec = 'Field',by = 'Stage')
preplant_soil_N_effects = contrast(preplant_soil_N_means, method = 'pairwise', adjust = "tukey")

summary(preplant_soil_N_effects)
cld(preplant_soil_N_means)
```
> Preplant soil N uptake different at all 3 sampling timepoints.

## Graphing preplant soil_N mean and SE

```{r}
preplant_soilN_graphing <- preplant %>% group_by(Field, Days) %>%  
  mutate(soil_N_se = sd(soil_N)/sqrt(3)) %>%   
  summarise(soil_N = mean(soil_N),
            soil_N_se = mean(soil_N_se))
```

```{r}
preplant_soil_N_graph <-
ggplot(preplant, aes(x=Days, y=soil_N, color=Field))+
  geom_point(data=preplant_soilN_graphing, size=2.5)+
  geom_smooth(method = lm, alpha=0)+
  scale_color_manual(values=c("#0072B2","#FFCC66"), name = "Field", labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  scale_x_continuous(name="Days after seeding", limits = c(0, 140), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  scale_y_continuous(name=expression("N uptake (kg N ha"^{-1}*")"), limits = c(0, 150), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  geom_errorbar(data=preplant_soilN_graphing, aes(ymin=soil_N-soil_N_se, ymax=soil_N+soil_N_se), width=3)+
  #geom_vline(xintercept = c(41, 50, 78, 84, 121, 136), linetype = "dashed", color = "black") +
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size=16))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+  
  theme(plot.title = element_text(hjust = 0.5, size = 15))+
  annotate(
  "text",
  x = c(46,81, 127),  # X-axis positions for annotations
  y = c(75,100, 120),  # Y-axis positions for annotations
  label = "*",
  size = 8,
  vjust = 0  # Adjust vertical position of asterisks
)+
  annotate(
    "text",
    x = c(80),
    y = c(140),
    label = expression(paste("Preplant N uptake"[soil])),
    size = 7,
    vjust = 0
  )
preplant_soil_N_graph

ggsave(preplant_soil_N_graph, filename = "preplant_soil_N_graph.png", height = 15, width = 20, units = "cm", dpi = 1000)
```


# Preplant Fert N

## Model selection and testing

```{r}
preplant_fert_N_model <- lmer(fertiliser_N~Field*Year*Stage+(1|Blk)+(1|Blk:Field), data = preplant)
anova(preplant_fert_N_model)
```
> Field appears not significant 


```{r}
preplant_fert_N_means = emmeans(preplant_fert_N_model,spec = 'Field',by = 'Stage')
preplant_fert_N_effects = contrast(preplant_fert_N_means, method = 'pairwise', adjust = "tukey")

summary(preplant_fert_N_effects)
cld(preplant_fert_N_means)
```

> Preplant Fert N appears to only be different at PI. 

## Graphing preplant fert_N mean and SE

```{r}
preplant_fertiliser_N_graphing <- preplant %>% group_by(Field, Days) %>%  
  mutate(fertiliser_N_se = sd(fertiliser_N)/sqrt(3)) %>%   
  summarise(fertiliser_N = mean(fertiliser_N),
            fertiliser_N_se = mean(fertiliser_N_se))
```

```{r}
preplant_fertiliser_N_graph <-
ggplot(preplant, aes(x=Days, y=fertiliser_N, color=Field))+
  geom_point(data=preplant_fertiliser_N_graphing, size=2.5)+
  geom_smooth(method = lm, alpha=0)+
  scale_color_manual(values=c("#0072B2","#FFCC66"), name = "Field", labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  scale_x_continuous(name="Days after seeding", limits = c(0, 140), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  scale_y_continuous(name=expression("N uptake (kg N ha"^{-1}*")"), limits = c(0, 150), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  geom_errorbar(data=preplant_fertiliser_N_graphing, aes(ymin=fertiliser_N-fertiliser_N_se, ymax=fertiliser_N+fertiliser_N_se), width=3,position=position_dodge(0.5))+
  #geom_vline(xintercept = c(41, 50, 78, 84, 121, 136), linetype = "dashed", color = "black") +
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size=16))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  theme(plot.title = element_text(hjust = 0.5, size = 15))+
  annotate(
    "text",
    x = c(80),
    y = c(140),
    label = expression(paste("Preplant N uptake"[fertilizer])),
    size = 7,
    vjust = 0
  )+
  annotate("text",
  x = c(46),  # X-axis positions for annotations
  y = c(50),  # Y-axis positions for annotations
  label = "*",
  size = 8,
  vjust = 0  # Adjust vertical position of asterisks
)


preplant_fertiliser_N_graph



ggsave(preplant_fertiliser_N_graph, filename = "preplant_fertiliser_N_graph.png", height = 15, width = 20, units = "cm", dpi=1000)
```

# Topdress Soil N

## Model selection and testing

```{r}
topdress_soil_N_model <- lmer(soil_N~Field*Year*Stage+(1|Blk)+(1|Blk:Field), data = topdress)
anova(topdress_soil_N_model)
```

> Field has strong effect on topdress soil N uptake. 

```{r}
topdress_soil_N_means = emmeans(topdress_soil_N_model,spec = 'Field',by = 'Stage')
topdress_soil_N_effects = contrast(topdress_soil_N_means, method = 'pairwise', adjust = "tukey")

summary(topdress_soil_N_effects)
cld(topdress_soil_N_means)
```


## Graphing topdress soil_N mean and SE

```{r}
topdress_soil_N_graphing <- topdress %>% group_by(Field, Days) %>%  
  mutate(soil_N_se = sd(soil_N)/sqrt(3)) %>%   
  summarise(soil_N = mean(soil_N),
            soil_N_se = mean(soil_N_se))
```

```{r}
topdress_soil_N_graph <-
ggplot(topdress, aes(x=Days, y=soil_N, color=Field))+
  geom_point(data=topdress_soil_N_graphing, size=2.5)+
  geom_smooth(method = lm, alpha=0)+
  scale_color_manual(values=c("#0072B2","#FFCC66"), name = "Field", labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  scale_x_continuous(name="Days after seeding", limits = c(60, 140), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  scale_y_continuous(name=expression("N uptake (kg N ha"^{-1}*")"), limits = c(0, 150), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  geom_errorbar(data=topdress_soil_N_graphing, aes(ymin=soil_N-soil_N_se, ymax=soil_N+soil_N_se), width=3,position=position_dodge(00))+
  #geom_vline(xintercept = c(78, 84, 121, 136), linetype = "dashed", color = "black") +
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size=16))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  theme(plot.title = element_text(hjust = 0.5, size = 15))+
  annotate(
  "text",
  x = c(81, 127),  # X-axis positions for annotations
  y = c(87, 110),  # Y-axis positions for annotations
  label = "*",
  size = 8,
  vjust = 0  # Adjust vertical position of asterisks
)+
    annotate(
    "text",
    x = c(100),
    y = c(140),
    label = expression(paste("Topdress N uptake"[soil])),
    size = 7,
    vjust = 0
  )


topdress_soil_N_graph

ggsave(topdress_soil_N_graph, filename = "topdress_soil_N_graph.png", height = 15, width = 20, units = "cm", dpi = 1000)
```

# Topdress Fert N

## Model selection and testing

```{r}
topdress_fert_N_model <- lmer(fertiliser_N~Field*Year*Stage+(1|Blk)+(1|Blk:Field), data = topdress)
anova(topdress_fert_N_model)
```

> Field appears significant for preplant fert N. 

```{r}
topdress_fert_N_means = emmeans(topdress_fert_N_model,spec = 'Field',by = 'Stage')
topdress_fert_N_effects = contrast(topdress_fert_N_means, method = 'pairwise', adjust = "tukey")

summary(topdress_fert_N_effects)
cld(topdress_fert_N_means)
```

> Preplant fert N uptake different at heading and maturity. 

## Graphing topdress fert_N mean and SE

```{r}
topdress_fertiliser_N_graphing <- topdress %>% group_by(Field, Days) %>%  
  mutate(fertiliser_N_se = sd(fertiliser_N)/sqrt(3)) %>%   
  summarise(fertiliser_N = mean(fertiliser_N),
            fertiliser_N_se = mean(fertiliser_N_se))
```

```{r}
topdress_fertiliser_N_graph <-
ggplot(topdress, aes(x=Days, y=fertiliser_N, color=Field))+
  geom_point(data=topdress_fertiliser_N_graphing, size=2.5)+
  geom_smooth(method = lm, alpha=0)+
  scale_color_manual(values=c("#0072B2","#FFCC66"), name = "Field", labels = c("Continuous rice (CR)", "Fallow rice (FR)"))+
  scale_x_continuous(name="Days after seeding", limits = c(60, 140), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  scale_y_continuous(name=expression("N uptake (kg N ha"^{-1}*")"), limits = c(0, 150), expand = c(0, 0), breaks = seq(0, 150, by = 20))+
  geom_errorbar(data=topdress_fertiliser_N_graphing, aes(ymin=fertiliser_N-fertiliser_N_se, ymax=fertiliser_N+fertiliser_N_se), width=3,position=position_dodge(0.5))+
  #geom_vline(xintercept = c(78, 84, 121, 136), linetype = "dashed", color = "black") +
  theme_classic()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size=16))+
  theme(legend.text = element_text(size = 12),legend.title = element_text(size = 14))+
  #ggtitle("Topdress fertilizer N uptake")+
  theme(plot.title = element_text(hjust = 0.5, size = 15))+
  annotate(
  "text",
  x = c(81, 127),  # X-axis positions for annotations
  y = c(15, 15),  # Y-axis positions for annotations
  label = "*",
  size = 8,
  vjust = 0  # Adjust vertical position of asterisks
)+
   annotate(
    "text",
    x = c(100),
    y = c(140),
    label = expression(paste("Topdress N uptake"[fertilizer])),
    size = 7,
    vjust = 0
  )

topdress_fertiliser_N_graph

ggsave(topdress_fertiliser_N_graph, filename = "topdress_fertiliser_N_graph.png", height = 15, width = 20, units = "cm")
```

# Total N uptake

```{r}
all_data$total_N <- all_data$fertiliser_N + all_data$soil_N

all_data %>% group_by(Topdress, Stage, Field) %>% summarise(total_N = mean(total_N))
```

```{r}
preplant_all <- cbind(preplant_soilN_graphing,
                   preplant_fertiliser_N_graphing %>% ungroup()%>% select(-Days, -Field) %>% mutate(Topdress = "Preplant"))

topdress_all <- cbind(topdress_soil_N_graphing,
                   topdress_fertiliser_N_graphing %>% ungroup()%>% select(-Days, -Field) %>% mutate(Topdress = "Topdress"))

for_excel <- rbind(preplant_all, topdress_all)

write_xlsx(for_excel, "fert_soil_N_by_trt_and_topdress.xlsx")
```

